---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# GEO712PACKAGE

The GEO712PACKAGE is a test R package created for the GEO712 course at McMaster University (fall term of 2023). In this R package we share the processed data of a multiple regression model. We use a data set from Kaggle, named "House Sales in King County, USA", to predict house price using a regression model. The data set contains house sales prices for King County, which includes Seattle. It includes homes sold between May 2014 and May 2015, with 21,613 registers and 21 variables regarding the houses sold.

## Installation

You can install the development version of GEO712PACKAGE from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("dias-bruno/GEO712PACKAGE")
```

## Example

This is a basic example which shows you how to access the data files. First, we'll need call the library "GEO712PACKAGE:

```{r example}
library(GEO712PACKAGE)
```

Accessing the data:

```{r}
data(df)
houses <- df
```

Visualizing the summary of the data:

```{r}
summary(houses)
```

Now, we already have access to the data and we can create some graphs about the data. The next chunk shows the house prices predicted compared to the real values:

```{r}
plot(x = houses$predicted_values, y = houses$price, ylab = "House Price ($)", xlab = "Price ($) Predicted", col='red', cex = 0.5, pch = 19) + abline(lm(houses$price ~ houses$predicted_values), lwd=2, col='blue')
```
